When, why, and how to avoid, prefer, or tolerate inbreeding

### 2018-03-03 presentation/inbreeding.Rmd
Add more images

- Inbreeding avoidance is predicted given a sufficiently high threshold of inbreeding depression [1,2]

- Inbreeding depression thresholds differ between females and males causing sexual conflict [1,2]

@@

- Inbreeding avoidance is predicted given a sufficiently high threshold of inbreeding depression [1,2]

- Inbreeding depression thresholds differ between females and males causing sexual conflict [1,2]

- **Above** thresholds we predict inbreeding avoidance

- **Below** thresholds we predict inbreeding preference (females) or tolerance (males) [3,4]

@@

- Inbreeding tolerance increases with lower encounter rates [1].

@@

- Inbreeding tolerance increases with lower encounter rates [1].

- Inbreeding avoidance or preference can affect dispersal [2].

@@

- Inbreeding tolerance increases with lower encounter rates [1].

- Inbreeding avoidance or preference can affect dispersal [2].

- Thresholds change given mating decisions of avoided relatives [3].

@@

- Inbreeding tolerance increases with lower encounter rates [1].

- Inbreeding avoidance or preference can affect dispersal [2].

- Thresholds change given mating decisions of avoided relatives [3].

- Thresholds change if parental investment can be modulated [4].

@@

- Inbreeding tolerance increases with lower encounter rates [1].

- Inbreeding avoidance or preference can affect dispersal [2].

- Thresholds change given mating decisions of avoided relatives [3].

- Thresholds change if parental investment can be modulated [4].

- Thresholds change if there are male opportunity costs [5]

@@

- Focuses on selection to the exclusion of drift, mutation, immigration

@@

- Focuses on selection to the exclusion of drift, mutation, immigration

- Does not consider realistic distributions of kinship

@@

- Focuses on selection to the exclusion of drift, mutation, immigration

- Does not consider realistic distributions of kinship

- Does not consider complex genetic architecture

@@

- Focuses on selection to the exclusion of drift, mutation, immigration

- Does not consider realistic distributions of kinship

- Does not consider complex genetic architecture

- Does not account for post-copulatory inbreeding avoidance or preference

@@

- Models discrete individual females and males in a finite population

- Explicitly tracks pedigree to calculate kinship between individuals

- Account for complex genetic architecture underlying phenotypes

- Allows for the evolution of pre- and post-copulatory inbreeding, and polyandry

@@

- **When should we predict selection for a alleles causing pre-copulatory inbreeding avoidance or inbreeding preference? [1]**

- **Why should we predict females will mate multiply to adjust the degree to which they inbreed? [2]**

- **How should we predict females to avoid inbreeding (pre- versus post-copulatory avoidance)? [3]**

@@

@@

- Fact one of simulation
- Fact two

@@

Histograms -- females and males, versus neutral, add costs.

Emphasis on stochasticity

Population size important

Selection coefficients, inbreeding coefficients, and delta on a half-sib or cousin

Agreement with an infinite alleles model with 10 loci.

>>>>

Inbreeding avoidance is predicted given a sufficiently high threshold of inbreeding depression [1,2]

Inbreeding depression thresholds differ between females and males causing sexual conflict [1,2]

@@

Inbreeding avoidance is predicted given a sufficiently high threshold of inbreeding depression [1,2]

Inbreeding depression thresholds differ between females and males causing sexual conflict [1,2]

**Above** thresholds we predict inbreeding avoidance

**Below** thresholds we predict inbreeding preference (females) or tolerance (males) [3,4]

@@

Inbreeding tolerance increases with lower encounter rates [1].

@@

Inbreeding tolerance increases with lower encounter rates [1].

Inbreeding avoidance or preference can affect dispersal [2].

@@

Inbreeding tolerance increases with lower encounter rates [1].

Inbreeding avoidance or preference can affect dispersal [2].

Thresholds change given mating decisions of avoided relatives [3].

@@

Inbreeding tolerance increases with lower encounter rates [1].

Inbreeding avoidance or preference can affect dispersal [2].

Thresholds change given mating decisions of avoided relatives [3].

Thresholds change if parental investment can be modulated [4].

@@

Inbreeding tolerance increases with lower encounter rates [1].

Inbreeding avoidance or preference can affect dispersal [2].

Thresholds change given mating decisions of avoided relatives [3].

Thresholds change if parental investment can be modulated [4].

Thresholds change if there are male opportunity costs [5]

@@

Focuses on selection to the exclusion of drift, mutation, immigration

@@

Focuses on selection to the exclusion of drift, mutation, immigration

Does not consider realistic distributions of kinship

@@

Focuses on selection to the exclusion of drift, mutation, immigration

Does not consider realistic distributions of kinship

Does not consider complex genetic architecture

@@

Focuses on selection to the exclusion of drift, mutation, immigration

Does not consider realistic distributions of kinship

Does not consider complex genetic architecture

Does not account for post-copulatory inbreeding avoidance or preference

@@

Models discrete individual females and males in a finite population

Explicitly tracks pedigree to calculate kinship between individuals

Account for complex genetic architecture underlying phenotypes

Allows for the evolution of pre- and post-copulatory inbreeding, and polyandry

@@

**When should we predict selection for a alleles causing pre-copulatory inbreeding avoidance or inbreeding preference? [1]**

**Why should we predict females will mate multiply to adjust the degree to which they inbreed? [2]**

**How should we predict females to avoid inbreeding (pre- versus post-copulatory avoidance)? [3]**

@@

## Mutant allele underlying inbreeding avoidance {.smaller}

Compared frequency of a mutant allele underlying inbreeding avoidance to a mutant neutral allele after 3000 generations.

**100 simulation replicates across multiple parameter values:**

- Selection against deleterious recessives
- Dominance coefficient of deleterious recessives
- Direct selection against inbreeding avoidance
- Offspring produced per female
- Maximum mates per male
- Immigrants per generation
- Female and male carrying capacity
- Allele mutation rate (forward & back)
- Effect strength of inbreeding avoidance allele

@@

![](images/F_A.png)

1000 loci with deleterious recessive alleles (homozygotes 2\% survival cost)

Inbreeding depression slope $\beta \approx 1$

$\delta \approx 0.22$ for full siblings

$\delta \approx 0.12$ for half siblings

$\delta \approx 0.06$ for first cousins

![](images/F_A_c0pt01.png)

**Selection for inbreeding avoidance is no longer predicted given a direct fitness cost of 1% for carrying the allele.**

![](images/F_A_alpha20.png)

Selection for inbreeding avoidance is no longer predicted given a direct fitness cost of 1% for carrying the allele.

**Selection for inbreeding avoidance increases when the *effect* of the allele is strengthened (i.e., relatives are avoided at higher probability)**

![](images/M_A.png)

Selection for inbreeding avoidance is no longer predicted given a direct fitness cost of 1% for carrying the allele.

Selection for inbreeding avoidance increases when the *effect* of the allele is strengthened (i.e., relatives are avoided at higher probability)

**Selection for inbreeding avoidance comparable when males rather than females chose mates** (*Note:* Males were mate-limited)

@@

Add effective population size plot to females before showing males

After briefly showing males, then move on to showing inbreeding preference

Note that inbreeding preference is predicted for females, not males

Move on to different genetic architecture where inbreeding avoidance/preference in same sim
